---
title: "R Language Basics and Data Importing"
subtitle: "PSYC 259: Principles of Data Science"
author: "John Franchak"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: xaringan-themer.css
    nature:
      ratio: 16:9
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_mono_accent(
  base_color = "#1c5253",
  header_font_google = google_font("Josefin Sans"),
  text_font_google   = google_font("Montserrat", "300", "300i"),
  code_font_google   = google_font("Fira Mono"),
  base_font_size = "20px",
  text_font_size = "1.5rem",
  header_h1_font_size = "3rem",
  header_h2_font_size = "2.5rem",
  header_h3_font_size = "2rem",
)
```
```{r, include = F}
# This is the recommended set up for flipbooks
# you might think about setting cache to TRUE as you gain practice --- building flipbooks from scratch can be time consuming
knitr::opts_chunk$set(fig.width = 6, message = FALSE, warning = FALSE, comment = "", cache = F)
library(flipbookr)
library(tidyverse)
```



# R Language basics

- Calculation and assignment 

- Calling functions 

- Storing data in tibbles 
 
- Extending the R language with packages 


---
`r chunk_reveal("calc_vs_assign", break_type = "auto", widths = c(4, 1))`

```{r calc_vs_assign, include = FALSE}
# R can run simple calculations
1 + 1 
4^2

# Calculations aren't useful unless we put the results somewhere 
# The assignment operator stores the result into a variable

two <- 1 + 1
two

# Once assigned, variables can be modified by re-assigning a new value to them

two <- 1 + 3
two

# Variables can be removed from the workspace with rm
# After removing two, calling it again would lead to an error 
rm(two)

# Variables can be reused in expressions to calculate new variables/outputs
var1 <- 5
var2 <- 10
var3 <- var1 + var2

# When assigning variables, nothing prints to the console 
# Let's use the function print

print(var3)
```
---

`r chunk_reveal("functions", break_type = "auto", widths = c(4, 1))`

```{r functions, include = FALSE}
# Print and rm are functions that we use in R 
# We call functions by writing their name followed   
# by a list of arguments in parentheses

abs(-1)
mean(var1, var2)
sum(var1, var2, var3)

# c is a function that combines values together
my_vars <- c(var1, var2, var3)
my_vars

# RStudio has built-in help for every function
?c

```
---
class: inverse, center, middle

# Next up:  cycle through different lines of code





<!-- get flipbook that rotates through lines, my_fill -->
---

`r chunk_reveal("my_fill", break_type = "rotate", widths = c(1,1), title = "### cycle through lines of code marked #ROTATE")`

```{r my_fill, eval = F, echo = F}
ggplot(data = cars) +
  aes(x = speed) +
  aes(y = dist) +
  geom_point(size = 8,
             shape = 21,
             alpha = .9,
             color = "snow") +
  aes(fill = speed) +
  scale_fill_viridis_c(option = "magma") + #ROTATE
  scale_fill_viridis_c(option = "cividis") + #ROTATE
  scale_fill_viridis_c(option = "plasma") #ROTATE
```


---
class: inverse, center, middle

# Next up:  look at multiple realizations of same code





<!-- get flipbook that shows multiple realizations of the same code -->
---

`r chunk_reveal("cars_multi", break_type = 5, title = "### see multiple realizations of code")`

```{r cars_multi, include = FALSE}
cars %>% 
  sample_frac(size = 1, replace = TRUE) %>%      
  ggplot() +              
  aes(x = speed) +
  aes(y = dist) +
  geom_count(
    alpha = .7,
    color = "blue",
    size = 4
    ) + 
  geom_smooth(method = lm, se = FALSE) +
  coord_cartesian(xlim = range(cars$speed),
                  ylim = range(cars$dist)) +
  theme(legend.position = c(.9, .2))
```


---

`r chunk_reveal("test", left_assign = "detect")`

```{r test, include=FALSE}
# hello
jam <- "hi"
```



<!-- adjust font size in this css code chunk, currently 80 -->

```{css, eval = TRUE, echo = FALSE}
.remark-code{line-height: 1.5; font-size: 70%}

@media print {
  .has-continuation {
    display: block;
  }
}

code.r.hljs.remark-code{
  position: relative;
  overflow-x: hidden;
}

}
```



---

# Notes: 

- This minimal flipbook assumes you are online when you build and view.  It is going to find the most recent remark.js info online.

- xaringan/remark.js slideshows are not self contained, single file slide shows.  Make sure you ship all the associated files that are created to display your flipbooks/slideshows.  I like to push .html and associated files to github repository and use Github Pages to host for beyond local sharing. :-)

